---
title: "CodeBook.md"
author: "Ted Herring"
date: "March 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Class Project
### Coursera, Getting and Cleaning Data, Week 4
### Johns Hopkins, Bloomberg School of Public Health


###Subjects Performing Activities
Name: Subject

There are 30 total subjects coded with numbers 1 to 30. Each subject performed a series of activities up to a maximum of 6 (six). A subject identification code list is not provided.


###Activities Performed
Name: Activity

There are 6 total activities that were to be performed by the subjects.

        1 WALKING
        2 WALKING_UPSTAIRS
        3 WALKING_DOWNSTAIRS
        4 SITTING
        5 STANDING
        6 LAYING



###MEASUREMENT VARIABLES^1^
The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

The set of variables that were estimated from these signals are: 
* mean: Mean value
* std: Standard deviation


####Naming Convention
The variable names of the measurements taken follow a pattern that describes the measurement.

For example: tBodyAcc.mean...X

```{r eval = FALSE}
{ <x><descriptor><sep><axis> }

```


```{r eval = FALSE}
<x> 

```
* t - time
* f - frequency


```{r eval = FALSE}
<sep>

```
* ...  if axis is specified
* ..   if axis is unspecified
The separator was originally '()-'. Is easier for humans to read.

```{r eval = FALSE}
<axis>

```

* x: x-axis
* y: y-axis
* z: z-axis

```{r eval = FALSE}
<descriptor>

```


  # | descriptor
-----|-----------
     | 
1    | BodyAcc-XYZ
2    | GravityAcc-XYZ
3    | BodyAccJerk-XYZ
4    | BodyGyro-XYZ
5 | BodyGyroJerk-XYZ
6 |BodyAccMag
7 |GravityAccMag
8 |BodyAccJerkMag
9 |BodyGyroMag
10 |BodyGyroJerkMag
11 |BodyAcc-XYZ
12 |BodyAccJerk-XYZ
13 |BodyGyro-XYZ
14 |BodyAccMag
15 |BodyAccJerkMag
16 |BodyGyroMag
17 |BodyGyroJerkMag





####COMPLETE LIST OF VARIABLES^1^

| variable name
|-----
|tBodyAcc.mean...X
|tBodyAcc.mean...Y
|tBodyAcc.mean...Z
|tBodyAcc.std...X
|tBodyAcc.std...Y
|tBodyAcc.std...Z
|tGravityAcc.mean...X
|tGravityAcc.mean...Y
|tGravityAcc.mean...Z
|tGravityAcc.std...X
|tGravityAcc.std...Y
|tGravityAcc.std...Z
|tBodyAccJerk.mean...X
|tBodyAccJerk.mean...Y
|tBodyAccJerk.mean...Z
|tBodyAccJerk.std...X
|tBodyAccJerk.std...Y
|tBodyAccJerk.std...Z
|tBodyGyro.mean...X
|tBodyGyro.mean...Y
|tBodyGyro.mean...Z
|tBodyGyro.std...X
|tBodyGyro.std...Y
|tBodyGyro.std...Z
|tBodyGyroJerk.mean...X
|tBodyGyroJerk.mean...Y
|tBodyGyroJerk.mean...Z
|tBodyGyroJerk.std...X
|tBodyGyroJerk.std...Y
|tBodyGyroJerk.std...Z
|tBodyAccMag.mean..
|tBodyAccMag.std..
|tGravityAccMag.mean..
|tGravityAccMag.std..
|tBodyAccJerkMag.mean..
|tBodyAccJerkMag.std..
|tBodyGyroMag.mean..
|tBodyGyroMag.std..
|tBodyGyroJerkMag.mean..
|tBodyGyroJerkMag.std..
|fBodyAcc.mean...X
|fBodyAcc.mean...Y
|fBodyAcc.mean...Z
|fBodyAcc.std...X
|fBodyAcc.std...Y
|fBodyAcc.std...Z
|fBodyAcc.meanFreq...X
|fBodyAcc.meanFreq...Y
|fBodyAcc.meanFreq...Z
|fBodyAccJerk.mean...X
|fBodyAccJerk.mean...Y
|fBodyAccJerk.mean...Z
|fBodyAccJerk.std...X
|fBodyAccJerk.std...Y
|fBodyAccJerk.std...Z
|fBodyAccJerk.meanFreq...X
|fBodyAccJerk.meanFreq...Y
|fBodyAccJerk.meanFreq...Z
|fBodyGyro.mean...X
|fBodyGyro.mean...Y
|fBodyGyro.mean...Z
|fBodyGyro.std...X
|fBodyGyro.std...Y
|fBodyGyro.std...Z
|fBodyGyro.meanFreq...X
|fBodyGyro.meanFreq...Y
|fBodyGyro.meanFreq...Z
|fBodyAccMag.mean..
|fBodyAccMag.std..
|fBodyAccMag.meanFreq..
|fBodyBodyAccJerkMag.mean..
|fBodyBodyAccJerkMag.std..
|fBodyBodyAccJerkMag.meanFreq..
|fBodyBodyGyroMag.mean..
|fBodyBodyGyroMag.std..
|fBodyBodyGyroMag.meanFreq..
|fBodyBodyGyroJerkMag.mean..
|fBodyBodyGyroJerkMag.std..
|fBodyBodyGyroJerkMag.meanFreq..


***
SOURCES AND METHODS

1. features_info.txt
2. R Documentation
3. how to Solve R Markdown knit Closure is not subsettable, https://stackoverflow.com/questions/37943824/how-to-solve-r-markdown-knit-closure-is-not-subsettable
4. Programmatically creating markdown tables in r with knitr, https://stackoverflow.com/questions/15488350/programmatically-creating-markdown-tables-in-r-with-knitr